\texttt{treats}: a modular \texttt{R} package for simulating trees and traits.

### 2023-11-22 paper/response_coloured.tex

I've added a \texttt{@return} section to the \texttt{treats},
\texttt{make.treats}, \texttt{make.bd.params}, \texttt{make.traits},
\texttt{make.modifiers} and \texttt{make.events} documentations as well
as the following section in the manual ``Getting started'' section:

\emph{\texttt{treats} will output either just a tree (class
\texttt{"phylo"}) if no traits were generated or a \texttt{"treats"}
object that contains both a \texttt{\$tree} (\texttt{"phylo"}) and a
\texttt{\$data} (\texttt{"matrix"}) component. Note that this
\texttt{"treats"} class is generalised to most outputs of the package
functions. This allows for a smoother handeling of the objects outputs
such as summarising the content of a \texttt{make.bd.params} output or
visualising a trait output from \texttt{make.traits}.}

\hypertarget{reviewer-2}{%
\section{Reviewer: 2}\label{reviewer-2}}

\textcolor{blue}{Thank you for the opportunity to review this manuscript, titled
``treats: a modular R package for simulating trees and traits''. This
manuscript describes and showcases a new R package that provides users
with an extensive toolbox to jointly simulate phylogenies and evolving
traits. While the example provided within the manuscript gives readers a
small taste of the flexibility of this toolbox, the online manual
(\url{http://tguillerme.github.io/treats.html}) showcases the true
extent of how useful this package will be for evolutionary biologists
far and wide. I really enjoyed reading the manuscript, and I believe it
will be suitable for publication in Methods in Ecology \& Evolution once
my concerns have been addressed.}

\textcolor{blue}{For context, for this review I read the supplied manuscript, tested the
functionality of the R package, and browsed the built-in documentation
and the online supplemental manual. I was unable to review the
functional code of the R package in time for the due date of this
review, but it appears to work properly from my testing. I have one
major concern and a handful of minor concerns. The major concern is that
the overall flow of the manuscript seems a little off. I think this
mostly stems from the worked example preceding the detailed overview of
the package's functions. I think if these two sections were swapped the
manuscript would flow much better. My minor concerns are included at the
end of this review in order of their appearance in the manuscript. I
look forward to reviewing a revised version of this manuscript, if
necessary.}

I have now changed the order of the structure of the manuscript with the
description coming first and the example coming second.

\textcolor{blue}{Best,}

\textcolor{blue}{William Gearty}

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Lines 27-31: I would enumerate this list.}
\end{itemize}

Done.

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Line 84: Geological time period names should be capitalized.}
\end{itemize}

Fixed.

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Figures 2-6: If possible, I think it would be valuable to show what
  code was used to produce these figures in the code snippets. As far as
  I can tell, the plotting is mostly done using functions from the
  package, so this would also be an opportunity to showcase this
  functionality.}
\end{itemize}

Done with the now new inclusion of the \texttt{dispRitreats} code that
streamlines the analysis.

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Figure 2: The caption should identify what the red line represents
  (the K-Pg boundary?).}
\end{itemize}

Done.

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Line 111: This may be personal preference, but when I ran this code
  chunk it output a whole bunch of text to the console. It appears to be
  a separate line of text for each replicate in the treats function. I
  believe the intent here is to show the user how much progress has been
  made? Perhaps some sort of progress bar would be more appropriate?}
\end{itemize}

I've now added a \texttt{verbose} option to the \texttt{treats} function
allowing to display the progress or not. Unfortunately the progress bar
suggestion, although comfortable to most users is not feasible
effectively with stochastic data: here the output indicates every re-ran
tree for each replicate. And because the number of runs varies depending
on the random seed, tracking the progress in terms of percentage will
require more computational time (i.e.~estimating how many re-run are
expected for the current seed before increase the progress bar).

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Lines 106-111: I think you should mention somewhere in this paragraph
  why you are running multiple replicates here (as opposed to the single
  replicate performed earlier in the worked example).}
\end{itemize}

I've added the following sentence at the end of the paragraph:
``\emph{Because of the stochasticity of the simulations, we will repeat
them 50 times (using \texttt{replicates\ =\ 50}) to generate a
distribution of possible simulated scenarios as opposed to a random
single one that could be idiosyncratic.}'' lines 200-203

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Line 111: I used the same seed as you, but I was unable to reproduce
  Figure 5 (looking at the R markdown file, I see that you use indices
  11 and 10 to make the plots). I think showing the exact code that is
  used here to make Figure 5 would be useful for readers that might be
  ``playing along at home''.}
\end{itemize}

I have streamlined and display the code to allow the users to test the
package while reading the paper.

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Lines 112-113: You should refer to Figure 6 somewhere in this
  paragraph. It might also be useful to include the code to perform this
  comparison, especially if the functionality is all contained within
  the treats package.}
\end{itemize}

Done. line 207.

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Line 194: This reference to the manual should have a link to the
  online manual (looks like the link is embedded, but you should also
  have it typed out). I also noticed that the documentation for the
  treats function lists the wrong URL for the online manual.}
\end{itemize}

I've added the URL spelt out.

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Lines 202-203: Just wanted to note that I really love this idea! I
  think collecting a library of modules will be very useful to users!}
\end{itemize}

As mentioned before, I was not sure how to effectively implement that so
I made a
\href{https://github.com/TGuillerme/treats/issues?q=is\%3Aopen+is\%3Aissue+label\%3A\%22simulation+template\%22}{simulation
templates} issues category on github for now.

\begin{itemize}
\tightlist
\item
  \textcolor{blue}{Line 205: This embedded link does not work. I think it should be
  something like:
  \url{https://github.com/TGuillerme/treats/tree/master/paper}.}
\end{itemize}

Fixed.

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-fossilsim}{}%
Barido-Sottani, Joëlle, Walker Pett, Joseph E O'Reilly, and Rachel CM
Warnock. 2019. ``FossilSim: An R Package for Simulating Fossil
Occurrence Data Under Mechanistic Models of Preservation and Recovery.''
\emph{Methods in Ecology and Evolution} 10 (6): 835--40.

\leavevmode\hypertarget{ref-rpanda}{}%
Morlon, Hélène, Eric Lewitus, Fabien L Condamine, Marc Manceau, Julien
Clavel, and Jonathan Drury. 2016. ``RPANDA: An R Package for
Macroevolutionary Analyses on Phylogenetic Trees.'' \emph{Methods in
Ecology and Evolution} 7 (5): 589--97.

\leavevmode\hypertarget{ref-puttick2020complex}{}%
Puttick, Mark N, Thomas Guillerme, and Matthew A Wills. 2020. ``The
Complex Effects of Mass Extinctions on Morphological Disparity.''
\emph{Evolution} 74 (10): 2207--20.

\leavevmode\hypertarget{ref-paleobuddy}{}%
Rosario Petrucci, Bruno do, Matheus Januario, and Tiago Quental. 2022.
``Paleobuddy: An R Package for Flexible Simulations of Diversification
and Fossil Sampling.'' \emph{Methods in Ecology and Evolution} 13 (12):
2692--8.

\leavevmode\hypertarget{ref-treesim}{}%
Stadler, Tanja. 2011. ``Simulating Trees with a Fixed Number of Extant
Species.'' \emph{Systematic Biology} 60 (5): 676--84.
\end{cslreferences}

\end{document}
